---
title: "ASSIGNMENT 3 FINAL"
author: "jxckieee"
date: "2024-09-12"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

title: "ASSIGNMENT 3 FINAL"
author: "jxckieee"
date: "2024-09-10"
output:
  pdf_document: default
  html_document: default
---

view(flights)
?flights

<!-- Write your Markdown below this line -->

flights_of_new_york.Rmd

##Exercise 1

i.  336776 entries, 19 columns

ii. flight number

iii.arrival time is later than scheduled arrival time

iv. carrier, origin

##Exercise 2

```{r}
flights %>%
  select(year, month) 
```  
```{r}
flights%>%
  select(year:day)
```
##Exercise 3

```{r}
flights %>%
  arrange(air_time, distance)
```
##Exercise 4

```{r}
flights %>%
  arrange(air_time, distance) 
```
##Exercise 5 

```{r}
flights%>%
  arrange(desc(month))
```
##Exercise 6

```{r}
flights%>%
  mutate(
    average_speed = distance / (air_time / 60)
  )
```
new  column: average_speed

##Exercise 7

```{r}
flights%>%
  mutate(
    dep_time_hour = dep_time %/% 100,
    dep_time_minute = dep_time %% 100
  )
```
```{r}
flights%>%
  mutate(
    dep_time_hour = dep_time %/% 100,
    dep_time_minute = dep_time %% 100,
    dep_time_minutes_midnight = (dep_time_hour * 60) + dep_time_minute
  )
```

##Exercise 8

```{r}
flights%>%
  filter(
    arr_delay < 0
  )
```

```{r}
flights%>%
  filter(
    carrier == "AA"
  )
```
##Exercise 9

```{r}
flights%>%
  group_by(carrier) %>%
  summarize(
    average_arr_delay = mean(arr_delay, na.rm = TRUE)
  )
```

##Exercise 10

```{r}
x <- 2
x <- 2 + 2
flights_to_miami <- flights %>%
  filter(dest == "MIA")
```

##Exercise 11

```{r}
monthly_delays <- flights %>%
  group_by(month, carrier) %>%
  summarize(
    arrival_delay = mean(arr_delay, na.rm = TRUE), 
    .groups = "drop"
  ) %>%
  spread(carrier, arrival_delay) %>%
  select(-`9E`)
```

```{r}
qplot(x = month, y = UA, geom="line", data=monthly_delays)
```
```{r}
monthly_delays %>%
  pivot_longer(cols = AA:YV, names_to = "airlines", values_to = "delays")
```