---
title: "Assignment 4: Capital Bikeshare"
author: "Jacquelyn Yeboah"
date: "`09/16/2024`"
documentclass: article
geometry: margin=1in
fontsize: 11pt
output:
  pdf_document:
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    dev: pdf
    highlight: tango
  html_document:
    theme: default
    self_contained: true
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    smart: true
    dev: svg
---
<!-- Write your Markdown below this line -->

## Exercise 1
i. start_lng and end_lng

ii. rideable_type and member_casual

iii. each row in the dataset represents a single bike ride.

## Exercise 2
```{r}
library(dplyr)

bikeshare <- bikeshare %>%
  mutate(
    duration = as.numeric(difftime(ended_at, started_at, units = "secs"))
  ) 

```

```{r}
bikeshare %>%
  ggplot() +
  geom_point(aes(x = duration, y = min_distance)) +
  labs(
    title = "Scatter plot of Min Distance vs Duration",
    x = "Duration (seconds)",
    y = "Min Distance"
  )

```

```{r}
summary(bikeshare$duration)
mean(bikeshare$duration)  # To calculate the mean duration
median(bikeshare$duration)  # To calculate the median duration

```

## Exercise 3
```{r}
bikeshare %>%
  ggplot () +
  geom_histogram(mapping = aes(started_at)) +
    labs (duration  = "...", started_at = "...")
```

## Exercise 4
```{r}
bikeshare %>%
  ggplot() +
  geom_boxplot(mapping = aes(y = duration)) +
  labs(title = "Box Plot of Trip Duration", y = "Duration (seconds)")
```
## Exercise 5
```{r}
library(ggplot2)

# Sample data
data <- data.frame(
  group = rep(c("A", "B", "C"), each = 100),
  value = c(rnorm(100), rnorm(100, mean = 2), rnorm(100, mean = 1.5))
)

# Create a violin plot
ggplot(data, aes(x = group, y = value)) +
  geom_violin()

```
```{r}
library(ggplot2)

# Create a violin plot
ggplot(bikeshare, aes(x = "start_lat", y = x)) +
  geom_violin() +
  xlab("")

```

Chat GPT was used to modify a violin plot, generated for this exercise.
## Exercise 6
```{r}
bikeshare <- bikeshare %>%
  mutate(
    min_distance = distHaversine(
      cbind(start_lng, start_lat), 
      cbind(end_lng, end_lat)
      )
      )

```


## Exercise 7
```{r}
library(ggplot2)

# Histogram
ggplot(bikeshare, aes(x = min_distance)) +
  geom_histogram(binwidth = 500, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Minimum Distance", x = "Minimum Distance (meters)", y = "Count") +
  theme_minimal()

```

```{r}
# Box plot
ggplot(bikeshare, aes(x = "duration", y = min_distance)) +
  geom_boxplot(fill = "lightgreen", color = "black") +
  labs(title = "Box Plot of Minimum Distance", y = "Minimum Distance (meters)") +
  theme_minimal()

```

```{r}
# Violin plot
ggplot(bikeshare, aes(x = "duration", y = min_distance)) +
  geom_violin(fill = "red", color = "black") +
  labs(title = "Violin Plot of Minimum Distance", y = "Minimum Distance (meters)") +
  theme_minimal()

``` 

## Exercise 8
```{r}
bikeshare %>%
  ggplot() +
  geom_point(aes(x = "duration", y = "min_distance")) +
  labs(
    title = "bikeshare",
    x = "duration",
    y = "min_distance"
  )

```

```{r}
bikeshare %>%
  ggplot(aes(x = duration, y = min_distance)) +
  geom_point(alpha = 0.5, color = "blue") + # Adjust transparency and color
  labs(
    title = "Covariation between Duration and Minimum Distance",
    x = "Duration (minutes)", # Adjust label as needed
    y = "Minimum Distance (km)" # Adjust label as needed
  ) +
  theme_minimal() # Optional: Apply a clean theme

```

```{r}
bikeshare %>%
# Example data (assuming 'duration' and 'distance' are the variables in question)
ggplot() +
  geom_bin2d (data = bikeshare_data, aes(x = "duration", y = "distance",(alpha = 0.3))) +
  labs(title = "2D Histogram of Rides", x = "Duration", y = "Distance")
``` 

```{r}
bikeshare %>%
  ggplot(aes(x = duration, y = min_distance)) +
  geom_point(alpha = 0.01, color = "blue") + # Set alpha to 0.01 for higher transparency
  labs(
    title = "Covariation between Duration and Minimum Distance",
    x = "Duration (minutes)",
    y = "Minimum Distance (km)"
  ) +
  theme_minimal()

``` 

```{r}
bikeshare %>%
  ggplot() +
  geom_point(aes(x = duration, y = min_distance)) +
  labs(
    title = "Scatter Plot of Min Distance vs Duration",
    x = "Duration",
    y = "Min Distance"
  )

```

```{r}
bikeshare %>%
  ggplot() +
  geom_point(aes(x = duration, y = min_distance), alpha = 0.01) +
  labs(
    title = "Scatter Plot of Min Distance vs Duration (with Transparency)",
    x = "Duration",
    y = "Min Distance"
  )

``` 
## Exercise 9
```{r}
bikeshare %>%
  summarize(
      mean = mean(duration),
      median = median(duration),
      standard_deviation = sd(duration),
      minimum = min(duration),
      maximum = max(duration)
  )
```

```{r}
bikeshare %>%
  group_by(member_casual) %>%
  summarize(
    mean = mean(duration),
    median = median(duration),
    standard_deviation = sd(duration),
    minimum = min(duration),
    maximum = max(duration)
  )

```

```{r}
bikeshare %>%
  group_by(member_casual) %>%
  summarize(
    n = n()
  )
```
## Exercise 10
```{r}
bikeshare %>%
  ggplot() +
  geom_histogram(mapping = aes(x = duration, fill = member_casual),
                 bins = 20, alpha = 0.3, position = "identity") +
  labs(title = "Ride Duration by Member Type", x = "Duration (seconds)", y = "Count")
```

```{r}
bikeshare %>%
  ggplot () +
  geom_violin(mapping = aes (x = member_casual, y = duration)) +
  labs(title = "Violin Plot of the Ride Duration by Member Type", y = "Duration (seconds)")
```

```{r}
bikeshare %>%
  ggplot() +
  geom_bin2d(mapping = aes(x = duration, y = min_distance)) +
  scale_fill_viridis_c(trans = "log") +
  facet_wrap(~ member_casual) +
  labs(title = "Heatmap of the Duration versus Minimum Distance by Member Type")
```

## Exercise 11
```{r}
bikeshare %>%
  ggplot() +
  geom_histogram(mapping = aes(x = duration, fill = member_casual),
                 bins = 20, alpha = 0.3, position = "identity") +
  scale_fill_manual(values = c("brown", "blue")) + #custom colors
  labs(title = "Ride Duration by Member Type",
       x = "Duration (log-scaled seconds)",
       y = "Count",
       fill = "Rider Type") +
  theme_minimal()
  
```
My Interpretation: 
According to the data shown, casual riders had a higher number of long duration rides versus member riders. Casual riders also take more leisured rides for recreation, compared to the member riders, who take shorter rides.

## Academic Integrity statement

ChatGPT was used to help complete this assignment in accordance with the AI policies in the syllabus.I also referred to GPT for an example on how to complete Exercise 11.

https://chat.openai.com/share/12e3c048-3b36-4c98-b191-d00313dea219 
https://chatgpt.com/share/66ef093c-68b0-800d-aca9-e2a0f07e2865 

