---
title: "Assignment 5: Modeling COVID-19"
author: "Jacquelyn Yeboah"
date: "09/22/2024"
documentclass: article
geometry: margin=1in
fontsize: 11pt
output:
  pdf_document:
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    dev: pdf
    highlight: tango
  html_document:
    theme: default
    self_contained: true
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    smart: true
    dev: svg
---

```{r setup, include = FALSE, echo=FALSE}
# DO NOT ALTER THIS CHUNK
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.width = 5,
  fig.asp = 0.618,
  dpi = 120,
  fig.align = "center",
  cache = FALSE
)
# Load required packages
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(scales))
covid_real <- read_csv("covid_counts.csv")
```

SIR stands for:

Susceptible (the number of people who can still catch the disease)
Infectious (the number of people who have currently got the disease, and can spread it to susceptible people)
Recovered (the number of people who have recovered from the disease, and who are immune to reinfection)


```{r}
balance_on_day_2 <- 1000 + 0.01 * 1000
```

```{r}
bank_balance = 1000

for(day in 2:365){
  bank_balance[day] <- bank_balance[day - 1] + 0.01 * bank_balance[day - 1]
}
```

## Exercise 1
```{r}
days_range <- 1:250

print(days_range)
```


## Exercise 2
```{r}
susceptible <- 1:300000000  #3*10^8 susceptible to the disease

print(susceptible)
```

```{r}
infectious <- 1:1 #1 person is infectious

print(infectious)
```

```{r}
recovered <- 0 # no recoveries yet

print(recovered)
```


## Exercise 3

```{r}
beta <- 0.25
gamma <- 0.125
N  <- 300000001
for (day in 2:250){
susceptible[day] <- susceptible[day-1] - (beta * susceptible[day-1] * infectious[day-1]) / N
}
```

```{r}
  for(day in 2:250){
    
      susceptible[day-1] <-300000001
      infectious[day-1] <- 1
      recovered [day-1] <- 0
}
```

```{r}
beta <- 0.25
gamma <- 0.125
N <- 300000001

#Results from day one
susceptible <- c(300000000)
infectious <- c(1)
recovered <- c(0)

# Loop through day 2 to day 250
for(day in 2:250){
  #Formula to calculate number of susceptible, infectious, and recovered individuals
  new_infections <- (beta*susceptible[day-1]* infectious[day-1]) / N
  new_recoveries <- gamma*infectious[day-1]
  
  susceptible[day] <- susceptible[day - 1] - new_infections
  infectious[day] <- infectious[day - 1] + new_infections - new_recoveries
  recovered[day] <- recovered[day - 1] + new_recoveries
  }

#Print the results:
tail(susceptible)
tail(infectious)
tail(recovered)
```

## Exercise 4
```{r}

#load library
library(tibble)

#days range vector:
days_range <- 1:250

#combining the vectors into a tibble:
covid_sim <- tibble(
  day = days_range,
  susceptible = susceptible,
  infectious = infectious,
  recovered = recovered
)
  
```

## Exercise 5
```{r}
read.csv("https://www.dropbox.com/scl/fi/a1w0hs834k9qtq8qgvbci/covidSim.csv?rlkey=8o4scy7hznfw2p5o1mtfpazpg&st=55ooo2ll&dl=1") -> covid_sim
covid_sim <- covid_sim[,2:5]
```

```{r}
# Load libraries
library(ggplot2)
library(tidyr)

# Reshape the dataframe into tidy format

  covid_sim %>%
  pivot_longer (cols = c(susceptible, infectious, recovered),
                names_to = "status",
                values_to = "count")

# View first few rows of the reshaped data
head(covid_sim)

ggplot() +
  geom_line(aes(x = day, y = c(susceptible, infectious, recovered))) +
  labs(
    title = "Simulated COVID-19 cases from SIR model",
    x = "Days in the USA",
    y = "Number of People",
    color = "case types") + 
    
  theme_minimal ()
```
## Exercise 6
```{r}
library(ggplot2)

# Plot Line Graph
ggplot() +
  geom_line(data = covid_sim, aes(x = days_range, y = infectious, color = "Simulated Cases"), size = 1) + # Simulated Cases
  geom_line(data = covid_real, aes(x = days_in_country, y = current_cases, color = "Real Cases"), size = 1) + # Real Cases
  labs(
    title = "Comparison of Simulated and Real COVID-19 Cases over time",
       x = "Days",
       y = "Number of Cases",
       color = "case types") + # Label for the legend
  theme_minimal ()

```

## Exercise 7
```{r}
library(ggplot2)

# Plot Line Graph
ggplot() +
  geom_line(data = covid_sim, aes(x = days_range, y = infectious, color = "Simulated Cases"), size = 1) + # Simulated Cases
  geom_line(data = covid_real, aes(x = days_in_country, y = current_cases, color = "Real Cases"), size = 1) + # Real Cases
  labs(
    title = "Comparison of Simulated and Real COVID-19 Cases over time",
       x = "Days",
       y = "Number of Cases",
       color = "case types") + # Label for the legend
  theme_minimal ()

scale_x_log10(labels=trans_format('log10', math_format(10^.y)))
scale_y_log10(labels=trans_format('log10', math_format(10^.x)))

```

## Exercise 8

codes from exercises 1-4

```{r}
infectious <- 1:1 #1 person is infectious

print(infectious)
```

```{r}
recovered <- 0 # no recoveries yet

print(recovered)
```

```{r}
beta_values <- seq(0.125, 0.5, by=0.025)
gamma <- 0.125
N  <- 300000001
for (day in 2:250){
susceptible[day] <- susceptible[day-1] - (beta * susceptible[day-1] * infectious[day-1]) / N
}
```

```{r}
  for(day in 2:250){
    
      susceptible[day-1] <-300000001
      infectious[day-1] <- 1
      recovered [day-1] <- 0
}
```

```{r}
# Define a range of beta values to test
beta_values <- seq(0.125, 0.5, by=0.025)
gamma <- 0.125
N <- 300000001

#Results from day one
susceptible <- c(300000000)
infectious <- c(1)
recovered <- c(0)

# Loop through day 2 to day 250
for(day in 2:250){
  #Formula to calculate number of susceptible, infectious, and recovered individuals
  new_infections <- (beta*susceptible[day-1]* infectious[day-1]) / N
  new_recoveries <- gamma*infectious[day-1]
  
  susceptible[day] <- susceptible[day - 1] - new_infections
  infectious[day] <- infectious[day - 1] + new_infections - new_recoveries
  recovered[day] <- recovered[day - 1] + new_recoveries
  }

#Print the results:
tail(susceptible)
tail(infectious)
tail(recovered)
```

```{r}

#load library
library(tibble)

#days range vector:
days_range <- 1:250

#combining the vectors into a tibble:
covid_sim <- tibble(
  day = days_range,
  susceptible = susceptible,
  infectious = infectious,
  recovered = recovered
)
  
```

code from exercise 7
```{r}

library(ggplot2)

# Plot Line Graph
ggplot() +
  geom_line(data = covid_sim, aes(x = days_range, y = infectious, color = "Simulated Cases"), size = 1) + # Simulated Cases
  geom_line(data = covid_real, aes(x = days_in_country, y = current_cases, color = "Real Cases"), size = 1) + # Real Cases
  labs(
    title = "Comparison of Simulated and Real COVID-19 Cases over time",
       x = "Days",
       y = "Number of Cases",
       color = "case types") + # Label for the legend
  theme_minimal ()

scale_x_log10(labels=trans_format('log10', math_format(10^.y)))
scale_y_log10(labels=trans_format('log10', math_format(10^.x)))
```
```{r}
days_range <- 1:250

print(days_range)
```

```{r}
susceptible <- 1:300000000  #3*10^8 susceptible to the disease

print(susceptible)
```

```{r}
# Setup initial values for SIR Model Parameters

N <- 1000000 # Total population
I <- 1 # Initial number of infected individuals
R <- 0 # Initial number of recovered individuals
s <- N - I - R # Initial number of susceptible individuals
gamma <- 0.1 # Recovery rate 

# Beta Values to test
beta_values <- seq(0.125, 0.5, by=0.025) # Testing beta between values

```

## Academic Integrity statement
I have used Chat GPT to assist me in Exercise 4,5 and 6. I have not copied directly from the app, however, here are my interactions with GPT:



